\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2013-04-16 doc/sw-system.tex
Finished appendix on probabilistic model

\sw is a web-based system that enables the discovery of strong
gravitational lenses in wide-field imaging surveys by arbitrarily large
numbers of people. Carefully produced color composite images are
displayed to the classifiers via a flexible interface, which records
their estimates of the positions of candidate lensed features. Simulated
lenses, and expert-classified non-lenses, are inserted into the stream
at random intervals; this training set is used to give the classifiers
feedback on their performance as well as estimate a dynamically-updated
probability for any given image containing a lens. High probability
systems are filtered into the Zooniverse \Talk discussion system; this
smaller sample is further classified and analysed by volunteers into a
final set. We analyze the classification of the training set itself, and
find that 1) $<$volunteers learn$>$... 2) $<$the filter works$>$...

@@

Our aim is to enable the construction of a sample of good lens
candidates. Since we aspire to making logical  decisions, we define a 
``good candidate'' as one which has a high posterior probability of
being a lens, given the data. The data in our case are the pixel values
of a colour image. However, we can greatly compress these complex,
noisy  datasets by asking our classifiers what they think about them. A
complete  classification in \sw consists of a set of Marker positions,
or none at all. The null set represents the statement from the
classifier that the image in question is $\saidNOT$ a lens, while the
placement of any  Markers indicates that the classifier considers this
image to contain a $\saidLENS$. We simplify the problem ...

>>>>

\sw is a web-based service
 that enables the discovery of strong
gravitational lenses in wide-field imaging surveys by arbitrarily large
numbers of people. Carefully produced color composite images are
displayed to the classifiers via a flexible interface, which records
their estimates of the positions of candidate lensed features. Simulated
lenses, and expert-classified non-lenses, are inserted into the stream
at random intervals; this training set is used to give the classifiers
feedback on their performance as well as estimate a dynamically-updated
probability for any given image containing a lens. High probability
systems are filtered into the Zooniverse \Talk discussion system; this
smaller sample is further classified and analysed by volunteers into a
final set. We analyze the classification of the training set itself, and
find that 1) $<$volunteers learn$>$... 2) $<$the filter works$>$...

@@

Our aim is to enable the construction of a sample of good lens
candidates. Since we aspire to making logical  decisions, we define a 
``good candidate'' as one which has a high posterior probability of
being a lens, given the data: $\pr(\LENS|\data)$. Our problem is to
approximate this probability. The data in our case are the pixel values
of a colour image. However, we can greatly compress these complex, noisy
sets of data by asking each classifier what they think about them. A
complete  classification in \sw consists of a set of Marker positions,
or none at all. The null set represents the statement from the
classifier that the image in question is $\saidNOT$ a lens, while the
placement of any  Markers indicates that the classifier considers this
image to contain a $\saidLENS$.

We simplify the problem by only using the Marker positions to assess
whether the classifier  correctly assigned the classification
$\saidLENS$ or $\saidNOT$ after viewing (blindly) a member of the
training set of subjects. We model each classifier's expertise with a
simple 2 by 2 ``confusion matrix,'' which is updated every time that
volunteer classifies a training subject using the following 
very simple estimate:
\be
  \pr(``X"|X,T) \approx \frac{N_{``X"}}{N_X}.
  \label{eq:app:fraction}
\ee
Here, $X$ stands for the true classification of the subject, \ie either
$\LENS$ or $\NOT$, while $``X''$ is the corresponding classification
made by the classifier on viewing the subject. $N_X$ is the number of
lenses the classifier has been shown, while $N_{``X"}$ is the number of 
times the classifier got their classifications right. $T$ stands for all
$N_{\LENS} + N_{\NOT}$ training data that the classifier has seen to
date.

The full confusion matrix of the $k^{\rm th}$ classifier is therefore:
\be
  \mathcal{M}^k = 
  \begin{bmatrix}
    \pr(\saidLENS|\NOT,T_k) & \pr(\saidLENS|\LENS,T_k) \\
    \pr(\saidNOT |\NOT,T_k) & \pr(\saidNOT |\LENS,T_k)
  \end{bmatrix}.
\ee
Note that these probabilities are normalized, \eg 
$\pr(\saidNOT |\NOT) = 1 - \pr(\saidLENS|\NOT)$.

Now, when this classifer views a test subject, 
it is this confusion matrix that will tell us how to update the
probability of that test subject being a $\LENS$. Let us suppose that
this subject has never been seen before: the prior probability that it
is (or contains) a lens is 
\be
  \pr(\LENS) = p_0
\ee
where we have to assign a value for $p_0$. In the CFHTLS, we might expect
something like 100 lenses in 430,000 images, so $p_0 = 2\times10^{-4}$
is a reasonable estimate. The classifier makes a classification $C_k$ 
($= \saidLENS or \saidNOT$).
We can apply Bayes' Theorem to update this 
prior probability into a posterior one using this new information:
\begin{align}
  & \pr(\LENS|C_k,T_k) = \\ \notag
  & \frac{\pr(C_k|\LENS,T_k)\cdot\pr(\LENS)}
{\left[ \pr(C_k|\LENS,T_k)\cdot\pr(\LENS) + \pr(C_k|\NOT,T_k)\cdot\pr(\NOT) \right]},
  \label{eq:app:first}
\end{align}
which can be evaluated numerically using the elements of the confusion
matrix.

As an example, suppose we have a {\it perfect} classifier, with
confusion matrix
\be
  \mathcal{M}^{\rm perfect} = 
  \begin{bmatrix}
    0.0 & 1.0 \\
    1.0 & 0.0
  \end{bmatrix},
\ee
who is given a fresh subject that actually is a $\LENS$. 
They submit $C = \saidLENS$. 
The resulting posterior probability for the classified
subject being a $LENS$ would be
\begin{align}
  \pr(\LENS|\saidLENS,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0
\end{align}
as expected. Meanwhile, an {\it obtuse} classifier, who always gets it
wrong and has the column-swapped confusion matrix, 
\be
  \mathcal{M}^{\rm obtuse} = 
  \begin{bmatrix}
    1.0 & 0.0 \\
    0.0 & 1.0
  \end{bmatrix},
\ee
would submit $C = \saidNOT$ for this subject, also leading to
\begin{align}
  \pr(\LENS|\saidNOT,T_k) &= \frac{1.0 \cdot p_0}
           {\left[ 1.0\cdot p_0 + 0.0\cdot(1 - p_0) \right]}
   &= 1.0.
\end{align}
This shows that  classifiers who are systematically wrong can 
nevertheless provide useful information. Indeed, the information content
of each classification can be estimated before the next classification
is made, and computed just from the confusion matrix. We define the 
expected information per classification, or expertise, in bits, as
\be
  \langle I_k \rangle = \frac{1}{2} 
                        \sum_{i} \left( \mathcal{M}^k_{i}  
              + \mathcal{M}^k_{i}\cdot\log_2{\mathcal{M}^k_i} \right),
  \label{eq:app:info}
\ee
where $i$ runs over all 4 elements of the confusion matrix 
$\mathcal{M}^k_i$. The second term in the sum is recognisable as a
negative entropy, while the first term and the factor of 1/2
account for the way the confusion matrix encodes the results for the
classification of two types of subject, and that the probabilities
associated with each type are normalised. \Eref{eq:app:info} gives the
required result, that 
both the {\it perfect} and
{\it obtuse} classifiers contribute 1 bit of information each, per
classification. 
The only classifier that contributes the minimum
information per classification, zero bits, is the {\it random} 
classifier,
whose confusion matrix elements are all 0.5.   We conservatively 
initialise all classifiers' confusion matrices to that of  the random
classifier, effectively assuming that all users are equally poorly
informed about what a $\LENS$ is, and what is $\NOT$. Plotting 
$\langle I_k \rangle$ as a function of time will illustrate the learning
process undergone by the $k^{\rm th}$ classifier.

Suppose the $k+1^{\rm th}$ classifier now submits a classification,
on this subject. We can generalise \Eref{eq:app:first} by replacing the
prior probability with the current posterior probability:
\begin{align}
  \pr(\LENS & |C_{k+1},T_{k+1},\data) = \\ \notag
  & \frac{1}{Z} \pr(C_{k+1}|\LENS,T_{k+1}) \cdot \pr(\LENS|\data) \\ \notag
{\rm where}\;\; Z = & \pr(C_{k+1}|\LENS,T_{k+1})\cdot\pr(\LENS|\data) \\ \notag
      & + \pr(C_{k+1}|\NOT,T_{k+1})\cdot\pr(\NOT|\data),
  \label{eq:app:update}
\end{align}
and $\data = \{C_k,T_k\}$ is the set of all previous
classifications, each given their classifier's training subjects.
$\pr(\LENS|\data)$ is the fundamental property of each test subject that
we are seeking. We track $\pr(\LENS|\data)$ as a function of time,
and by comparing it to a lower or upper thresholds, make decisions about
whether to retire the subject from the classification interface or
promote it in \Talk, respectively.
